Binning Strategies and Related Loss for Large Data

### 2015-04-24 Paper.Rnw
legend in colourbar. bit better, not sure that we can make it much better

<<BaseballBinnedScatters2X10,echo=F,include=T,eval=T,fig.width=12, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, warning=F, fig.show='hold',fig.cap="Binned Scatterplots for Games vs Strikeouts">>=  
 # raw frequencies plot
p1 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="Standard Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_gradient("Frequencies", low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.text=element_text(size=12) )
# # bin frequencies distribution hugely skewed
# qplot(binfreq, geom="density", data= binout1[[1]], color=I("blue")) + theme_bw()
# qplot(log(binfreq), geom="density", data= binout1[[1]], color=I("blue")) + theme_bw()

## quantile freq binning
binoutQuantFreqGroups <- freqBin(binout1, binType="quantile", ncolor=4, logCount=FALSE)
labels = sort(unique(binoutQuantFreqGroups[[1]]$freqlabel))
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p2 <- qplot(binxs, binys, geom="tile", fill=factor(freqlabel), data=binoutQuantFreqGroups[[1]],
            main="Quantile Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_manual("Frequencies", values=cols, 
                    guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                         title.position = "top"),
                    labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.text=element_text(size=12) )

## standard log freq binning
binoutLogFreqGroups <- freqBin(binout1, binType="standard", ncolor=4, logCount=TRUE)
binoutLogFreqGroups[[1]]$freqlabel <- factor(binoutLogFreqGroups[[1]]$freqlabel, 
                                             levels = unique(binoutLogFreqGroups[[1]]$freqlabel)[c(3,5,2,4,1)])
labels = levels(binoutLogFreqGroups[[1]]$freqlabel)
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)

p3 <- qplot(binxs, binys, geom="tile", fill=factor(freqlabel), data=binoutLogFreqGroups[[1]],
      main="Standard Log Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_manual("Log Frequencies", values=cols, 
                    guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                         title.position = "top"),
                    labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.text=element_text(size=12) )

>>>>

<<BaseballBinnedScatters2X10,echo=F,include=T,eval=T,fig.width=12, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, warning=F, fig.show='hold',fig.cap="Binned Scatterplots for Games vs Strikeouts">>=  
 # raw frequencies plot
p1 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="Standard Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_gradient("Frequencies", low="#56B1F7", high="#132B43",
                      guide=guide_colourbar( 
                                        legend.key.width = unit(1.5, "cm"),
                                             legend.text=element_text(size=12),
                                            label.hjust=0.5,  
                                         title.position = "top")) + 
  theme_bw() + theme(aspect.ratio=1, 
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.position="bottom")
# # bin frequencies distribution hugely skewed
# qplot(binfreq, geom="density", data= binout1[[1]], color=I("blue")) + theme_bw()
# qplot(log(binfreq), geom="density", data= binout1[[1]], color=I("blue")) + theme_bw()

## quantile freq binning
binoutQuantFreqGroups <- freqBin(binout1, binType="quantile", ncolor=4, logCount=FALSE)
labels = sort(unique(binoutQuantFreqGroups[[1]]$freqlabel))
upper <- as.numeric(gsub(".*,([-+]?[0-9]*.?[0-9]+([eE][-+]?[0-9]+)?).*","\\1", as.character(labels) ))
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p2 <- qplot(binxs, binys, geom="tile", fill=factor(freqlabel), data=binoutQuantFreqGroups[[1]],
            main="Quantile Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_manual("Frequencies (Quartiles)", values=cols, labels=upper,
#                    guide = guide_legend(label.position="bottom", label.hjust=0.5,  
#                                         title.position = "top"),
                    guide = guide_legend(label.position="bottom", label.hjust=1.1,  
                                         title.position = "top")) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1.5, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.text=element_text(size=12) )

## standard log freq binning
binoutLogFreqGroups <- freqBin(binout1, binType="standard", ncolor=4, logCount=TRUE)
binoutLogFreqGroups[[1]]$freqlabel <- factor(binoutLogFreqGroups[[1]]$freqlabel, 
                                             levels = unique(binoutLogFreqGroups[[1]]$freqlabel)[c(3,5,2,4,1)])
labels = levels(binoutLogFreqGroups[[1]]$freqlabel)
upper <- as.numeric(gsub(".*,([-+]?[0-9]*.?[0-9]+([eE][-+]?[0-9]+)?).*","\\1", as.character(labels) ))
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)

p3 <- qplot(binxs, binys, geom="tile", fill=factor(freqlabel), data=binoutLogFreqGroups[[1]],
      main="Standard Log Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_manual("Log Frequencies", values=cols, 
                    guide = guide_legend(label.position="bottom", label.hjust=1.1,  
                                         title.position = "top"),
                    labels=upper) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1.5, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.text=element_text(size=12) )
